Response to reviewers

### 2024-11-28 Manuscript/ecography_reponse_letter.tex
Update ecography_reponse_letter.tex

\begin{itemize} 
  \item \textbf{Capturing approximated mechanisms with different metrics} discussing the simulated result in more details. Lines 369-405
  \item \textbf{Empirical results} focusing on the empirical results. Lines 407-448
  \item \textbf{Caveats} discussing clearly the caveats of our approach. Lines 450-505.
  \item  \textbf{Box 2: ``so which metric do \textit\{I\} choose?''} proposing a didactic approach and reflections to hopefully easily help workers to choose their own metric. Lines 506-553.
  \item \textbf{Conclusion} to briefly wrap up our findings. Lines 555-572.
\end{itemize}

% \textbf{Item 4.}
%  Finally, the discussion section needs additional work to better address the key aspects of ``what,'' ``how,'' and ``why'' in functional analyses
% The introduction emphasizes applying these fundamental questions of ecological analysis to the study of functional diversity metrics
% However, this focus seems to be lost in the discussion, which lacks a strong connection to the introduction
% Given that this is the core of your study (as reflected in the title), it should be the primary focus of the discussion section.
%  I recommend revising the initial part of the introduction to better align it with your main question and hypotheses and to clearly convey the core message of your study
% For example, a more detailed interpretation of how your results on different metrics relate to this aspect of your framework would be helpful
% As I mentioned in my previous item, acknowledging the limitations of your methods is important, but it should be addressed in a separate paragraph following the interpretation of your results, rather than being the primary focus of the discussion.

% We have now linked the framework of ``what'', ``how'' and ``why'' more clearly in the discussion:
%  \textit{For example, say one is interested in looking at changes in functional diversity using a divergence metric calculated with the probability density method (e.g.~\textbackslash texttt\{TPD::REND\}).}
%  \textit{They can apply this metric to their trait space and get two relative metric scores of, says, 0.7 and 0.65 (pattern) for two levels of extinction (process).}
%  \textit{The interpretation of this score as a change in functional diversity will depend on the worker's assumption of which mechanism affected the trait space.}
%  \textit{In our simulations, if the mechanism approximated was the equalising one, the pattern can clearly be interpreted as a change in functional diversity, but if the mechanism is a facilitation one, the pattern can be meaningless (i.e.~not clearly distinguishable from a null mechanism - Fig
% \textbackslash ref\{Fig:simulation\_results\} shows this example by contrasting row 5, columns 1 and 2).}
%  \textit{In this made up example, the procedure is done in reverse order where the the pattern is specified first and then the process and the mechanism (``What process and mechanism are needed to interpret pattern X?'\,').}
%  \textit{In Box 2 we suggest an approach that specifies the process and the mechanism first (``how do I interpret pattern X under a specific process and mechanism'\,'; see Box 2).}
%  Lines @@@

% As well as in the new Box 2
% Lines @@@

% \textbf{Item 5.}
%  Ensure consistency in the use of terms such as ``elements of the model'' versus ``tips'' or ``species'', and adhere to a uniform citation format (e.g., using or omitting a comma before the year).

% We have now standardised terminology and provide a glossary in the supplementary materials (based on Mammola 2021).

% Overall, addressing these areas will significantly enhance the paper's impact, readability, and contribution to the field of functional ecology.

% Here are more detailed comments on the different sections of the paper.

% \textbf{Abstract}.
%  Line 3-6: Please rephrase to simplify and avoid brackets.
%  We changed the sentence to:
%  \textit{Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to various stressors
% In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism.}
%  Lines @@@

% Line 10-11: Please rephrase to avoid brackets.
%  We changed the sentence to:
%  \textit{We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.}
%  Lines @@@

% Line 14-15: If space requirements permit, please include more detailed information about your results
% The current description is too vague and does not effectively convey your main findings.
%  We've expanded a little on the results
% One point of keeping the vagueness in the abstract though was to not give the impression that we provide any prescriptive results but rather a broad understanding that the choice of metric is important.
%  \textit{We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).}
%  \textit{For example, the competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.}

% \textbf{Introduction.}
%  Line 25-28
% It appears to me that you are using the definition of a functional trait from Violle et al.~(2007) in Oikos
% Please include a citation for this reference if it is the case.
%  We've added the reference as suggested.

% Line 37
% I would avoid using the term ``new trait-focused ecology.'' While it's true that a more ``explicit'' approach to functional ecology has emerged over the past few decades, the foundation of functional analyses is rooted in the field of ecology itself
% For example, as you mentioned, Darwin's finches represent an early functional analysis centered on beak traits
% Although the specific term ``functional traits'' was less commonly used in the past, broader functional analyses were not uncommon, particularly in plant ecology (see, for example, Calow, P
% 1987
% ``Towards a Definition of Functional Ecology.'' Functional Ecology; and Duarte et al., 1995
% ``Comparative Functional Plant Ecology: Rationale and Potentials.'' Trends in Ecology \& Evolution).
%  This is a really good point
% We've removed the mention of ``new''.

% Line 73-74
% Please clarify how this is related to the previous question ``what'' and ``how''.
%  We've expanded the sentence to:
%  \textit{Mechanism description is then about linking the pattern and the process to some biological properties (sometimes referred to as ``rules'\,') and is at the core of answering the biological question at hand.}
%  Lines @@@

% Line 93-96
% Please explain why the chosen metric is not appropriate for the proposed example.
%  We've expanded the sentence to:
%  \textit{In such a scenario, measuring the occupied area (i.e.~\(km^{2}\)) of each population on each island (a pattern - ``what?'\,') will not be the most appropriate way to understand the potential competition between these populations because it might not give information on whether the two populations actually interact in some way - a very likely condition for competition to take place.}
%  Lines @@@

% Line 100
% I think it would be beneficial to more clearly address your research question here
% Is your goal simply to describe the patterns of functional diversity across different processes, or are you aiming to test how well different metrics capture the effects of various mechanisms? If it's the latter, how did you test this, and what are your expectations in relation to the null model?
%  We have now clarified what our goal and expectations are in the last paragraph of the introduction:
%  \textit{Our goal is to assess the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  Lines @@@

% Line 102
% Please refer to Table 1 together with figure 1, to provide more details on the ecological mechanisms.
%  We now refer to both the table and the figure.

% Line 103-106
% Please provide more information on why you chose to use a combination of empirical and simulated data
% From my understanding, your intention was to apply your framework within a more realistic context
% If that is the case, please clarify this point to strengthen your rationale.
%  We changed the sentence to:
%  \textit{Beyond simulations, we also used an empirical dataset of Hawaiian bird traits and used pre-historical and historical extinction events on Hawaiian islands (process) resulting in a trait space modified by extinctions (the pattern) capturing the effects of species extinction on species trait distributions (the mechanism).}
%  Lines @@@

% Line 106-111
% This content could be more suitable at the start of the discussion section
% Here you could maybe focus on outlining your hypotheses and predictions rather than presenting your conclusions
% By doing so, you establish a clearer context for interpreting your findings and guide the reader through the rationale behind your study.
%  We have now moved this sentence to the end of the conclusion (Lines @@@) to end the introduction on a clear description of our goals and expectations:

% \textit{Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure \textbackslash ref\{Fig:simulations\} and \textbackslash ref\{Tab:mechanisms\}).}
%  \textit{Our goal is to the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  \textit{To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based extinction events on Hawaiian islands before and after the year 1500 (process; how) and how different metrics (pattern; what) can lead to different interpretations of the data.}
%  \textit{We show that, that with a fixed process and mechanisms approximation, the choice of the statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data with the exact same data being sometimes interpreted in opposite ways.}

@@

% \textbf{Methods}.
%  Line 115
% Please provide more details of the elements of your model, as not all readers may be familiar with the Brownian Motion trait model.
%  We clarified the sentence as follows:
%  \textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e.~a pure birth speciation model) until reaching 200 elements in \textbackslash texttt\{R\} \textbackslash citep\{rcore\} using \textbackslash texttt\{treats\} \textbackslash citep\{guillerme2024treats\}.}
%  Lines @@@

% Line 131
% Please explain what the distance D is.
%  We clarified this point as follows:
%  \textit{Increasing density: by removing the required amount of pairs of points with a variable pairwise distance of at least \(D\) (i.e.~removing the \(n\) pairs of points that are at least \(D\) distance away from each other to remove the desired number of elements).}
%  Lines @@@

% Line 139-144
% I suggest breaking this section into two sentences and improve readability
% The current structure is challenging to follow.
%  We split this sentence in three for clarity:
%  \textit{Each bandwidth has an observed probability of sampling of \(n\) (proportional to the number of elements in that bandwidth) and is skewed to a probability of sampling of \(n 
% times m^{p}\)
% For bandwidths with a low density (low values of \(n\)) we used \(m\>0\) to increase their resampling probability, and the opposite for bandwidths with a high density (high values of \(n\); using \(m\<0\) to decrease sampling property)
% This effectively flattened the normal distribution of the data by some factor \(p\) (here we used \(p=3\)).}
%  Lines @@@

% Table 1
% Please provide references for the Equalizing and Facilitation mechanisms to offer clearer context for the focal ecological mechanisms studied.
%  We now added references to \textbackslash citep\{chesson2000mechanisms,barot2004mechanisms\} for the equalising mechanism and to \textbackslash citep\{bruno2003inclusion, danet2024species\} for the facilitation one.
%  Lines @@@

% Line 183
% Please provide the total number of the different metrics you used and refer to the table below (table 2).
%  We've now added a reference to table 2 and summarised the total amount of metrics.
%  Lines @@@

% Figure 1
% I believe adding a color legend to the figure would enhance its clarity.
%  We have now reworked figure 1 to be more didactic (especially for the results interpretation)
% We have not added the color legend to the plot itself (although we have added orange arrows to make it clearer) but in the caption to not overload an already busy figure.

% Figure 1
% Please include information in the figure caption explaining what the two types of graphs represent.
%  We added the following clarifications to the figure caption:
%  \textit{We then applied different stressors to the resulting trait space
% We represent the elements removed by the stressors in blue and the ones kept in the trait space in orange
% The first column illustrate the distribution of the elements in one dimension with the overlap between the removed (blue) and kept (orange) elements in brown (note we did not measure any metric on 1D spaces)
% The second column illustrate the same distributions in 2 dimensions
% We did not illustrate the distributions in 4 or 8 dimensions here
% We used the following stressors:}
%  Lines @@@

% Line 194-195
% Please improve clarity.
%  @@@TODO: rewrite following other reviewers comments (stats for results)

% Line 201-203
% Please simplify and improve clarity.
%  @@@TODO: rewrite following other reviewers comments (stats for results)

% Line 207-211
% Please provide a clearer explanation of the effect size and statistical significance of these model outcomes, similar to the explanation provided in lines 198-201.
%  @@@TODO: rewrite following other reviewers comments (stats for results)
%  @@@TODO: calculate effect sizes
%  We now added all the statistics results in the supplementary results from table 6 to 18.

% Line 221
% Please specify the added species.
%  @@@TODO: name of the two added marine species

% Table 2
% The aspect for the third row is missing.
%  We are not sure about what the reviewer meant here
% The tables are all formatted in LaTeX and we prefer to leave it to the production team to set it up as they like should the manuscript be accepted.

% \textbf{Results}.
%  Line 248-251
% Please provide more detail regarding these results (e.g., what the scaled metric scores are, the effect size, and the significance of data reduction caused by the stressor)
% Additionally, refer to the relevant section of Figure 2 (to explain how the observed trends in the figure support this assertion), and to the tables in supplementary material.
%  @@@TODO: add stats for results (following other reviewers comments)

% Line 252
% I would avoid the term ``naively''.
%  We removed this term.

% Line 254
% Please ensure that the use of italics follows the guidelines from the journal.
%  We did not find any specific guideline regarding emphasising words with italics but removed it nonetheless.

% Line 254-256
% Apply the same changes mentioned in the comment below to lines 248-251.
%  @@@TODO: add stats for results (following other reviewers comments)

% Line 260
% Please avoid using the exclamation mark.
%  We changed the punctuation to a dot.

% Line 262-269
% Please provide more detail regarding these results
% Additionally, refer to the table in supplementary material.
%  @@@TODO: add stats for results (following other reviewers comments)
%  We also now refer to the supplementary results:
%  \textit{For the empirical data, we see again a range of different possible interpretations depending on the metric describing the pattern (Figure \textbackslash ref\{Fig:empirical\_results\}, Supplementary table 5).}
%  Lines @@@

% Line 275-283
% This should be in the discussion section.
%  We moved this paragraph to the discussion on lines @@@.

% Figure 2
% Maybe I misunderstood, but I think that you are missing the adjusted R square from the figure.
%  We removed mention of the R2 (from a previous cluttered version of the figure) and now refer to it in the supplementaries in the figure caption.

% Figure 3
% Please provide more details in the figure caption.
%  Please include the results from the t-test in the supplementary material
% While the significance of the test is demonstrated through the gray plots in Figures 2 and 3, the relative values are also necessary.
%  The tables in the supplementary materials now contain the R2 values for both the Figures 2 and 3
% We have now clarified that in the figure captions:
%  \textit{The model results (including \(R^{2}\) and t-statistics) are available in supplementary table 4.}
%  Lines @@@ and Lines @@@

% \textbf{Discussion}.
%  Line 285-291
% This section effectively summarizes your main goal and key message
% However, a more detailed interpretation of your main findings is needed
% I believe that simply stating that ``different metrics capture different patterns (what), leading to inferences about different processes (how) and mechanisms (why)'' is insufficient
% I recommend focusing more on the performance of the various metrics to offer readers a clearer framework for selecting the appropriate metric for each mechanism
% Please refer to items 3 and 4 for further details.
%  Line 291-302
% I appreciate the detailed discussion of the potential limitations of your methodology
% However, in my opinion, this should follow a more in-depth interpretation of your results and their relation to the ``what'', ``how'', and ``why'' questions
% Consider placing the discussion of caveats and limitations in a separate paragraph, where you address the constraints of your framework.
%  Line 303-315
% Based on your preference, I recommend moving this section further into the discussion
% This will allow you to address the importance of selecting the appropriate metric in relation to the number of dimensions used for the trait space.
%  We have now re-orginased the discussion and clearly separated each subsection
% The first one now discusses the results (simulated and empirical), the second one focuses on the caveats of our approach
% And the third section is now a standalone box (box 2) discussing how we believe the what, how and why questions can help workers choose their functional diversity metric of interest based on their own data and question
% Hopefully these clear distinctions will make it easier to follow (or skip) for the reader.

% Line 313-316
% Please rephrase to improve readability.
%  We have now removed this sentence.

% Line 316-337
% I think this section should be simplified
% The discussion focuses too heavily on the limitations of the approach, making it overly methodological
% A greater emphasis on the ecological implications of the results would better convey your main message
% I recommend streamlining this section and, if preferred, moving part of it to the supplementary material, where you can address the implications of scaling your results with the null model.
%  We have now moved this section into the named ``caveats'' section
% Hopefully this will make it easier to skip for readers no interested in the methodological aspects of our work
% However, if the reviewer still thinks this section is too focused, we're happy to reduce it in a future version.

% Line 321-324
% When you refer to your findings, please provide a more explicit reference to the values you found and the relative graph to better guide the reader through your results and your interpretation.
%  We have now specified which part of figure 2 we were using to illustrate this point:
%  \textit{(e.g.~facilitation mechanism measured as divergence using the probability density method - not distinguishable from null when removing 80\textbackslash\% or 60\textbackslash\% of the data; Figure \textbackslash ref\{Fig:simulations\})}
%  Lines @@@

% Line 327
% Please replace `workers' with `ecologists,' as you are referring to an example from an ecological study.
%  We changed ``workers'' to ``researchers'' because not only ecologists are likely to work on such data/questions (e.g.~evolutionary biologists, palaeobiologists, etc.)

% Line 339-340
% Please simplify and increase readability.
%  We reworded this sentence as follows:
%  \textit{The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity since 1500 is not greater than expected by chance (at least for certain metrics; Figure \textbackslash ref\{Fig:empirical\_results\}) because one would expect that extinctions since 1500 target specific areas of functional diversity space (i.e.~birds with specific trait combinations).}
%  Lines @@@

>>>>

\begin{itemize} 
  \item \textbf{Capturing approximated mechanisms with different metrics} discussing the simulated result in more details. Lines 369-405
  \item \textbf{Empirical results} focusing on the empirical results. Lines 407-448
  \item \textbf{Caveats} discussing clearly the caveats of our approach. Lines 450-505.
  \item  \textbf{Box 2: ``so which metric do \textit{I} choose?''} proposing a didactic approach and reflections to hopefully easily help workers to choose their own metric. Lines 506-553.
  \item \textbf{Conclusion} to briefly wrap up our findings. Lines 555-572.
\end{itemize}

\textcolor{blue}{\textbf{Item 4.}}

\textcolor{blue}{Finally, the discussion section needs additional work to better address the key aspects of ``what,'' ``how,'' and ``why'' in functional analyses.
The introduction emphasizes applying these fundamental questions of ecological analysis to the study of functional diversity metrics.
However, this focus seems to be lost in the discussion, which lacks a strong connection to the introduction.
Given that this is the core of your study (as reflected in the title), it should be the primary focus of the discussion section.
I recommend revising the initial part of the introduction to better align it with your main question and hypotheses and to clearly convey the core message of your study.
For example, a more detailed interpretation of how your results on different metrics relate to this aspect of your framework would be helpful.
As I mentioned in my previous item, acknowledging the limitations of your methods is important, but it should be addressed in a separate paragraph following the interpretation of your results, rather than being the primary focus of the discussion.}

We have now linked the framework of ``what'', ``how'' and ``why'' more clearly in the discussion:

\textit{For example, say one is interested in looking at changes in functional diversity using a divergence metric calculated with the probability density method (e.g. \texttt{TPD::REND}).
They can apply this metric to their trait space and get two relative metric scores of, say, $0.7$ and $0.65$ (pattern) for two levels of extinction (process).
The interpretation of this score as a change in functional diversity will depend on the worker's assumption of which mechanism affected the trait space.
In our simulations, if the mechanism approximated was the equalising one, the pattern can clearly be interpreted as a change in functional diversity, but if the mechanism is a facilitation one, the pattern can be meaningless (i.e. not clearly distinguishable from a null mechanism - Fig. 2 shows this example by contrasting row 5, columns 1 and 2).
In this made up example, the procedure is done in reverse order where the the pattern is specified first and then the process and the mechanism (``What process and mechanism are needed to interpret pattern X?'').
In Box 2 we suggest an approach that specifies the process and the mechanism first (``how do I interpret pattern X under a specific process and mechanism''; see Box 2).}
Lines 385-398

As well as in the new Box 2 lines 506-553.

\textcolor{blue}{\textbf{Item 5.}}

\textcolor{blue}{Ensure consistency in the use of terms such as ``elements of the model'' versus ``tips'' or ``species'', and adhere to a uniform citation format (e.g., using or omitting a comma before the year).}

We have now standardised terminology and provide a glossary in the supplementary materials (table 1; based on Mammola 2021).

\textcolor{blue}{Overall, addressing these areas will significantly enhance the paper's impact, readability, and contribution to the field of functional ecology.}

\textcolor{blue}{Here are more detailed comments on the different sections of the paper.}

\textcolor{blue}{\textbf{Abstract}.}

\textcolor{blue}{Line 3-6: Please rephrase to simplify and avoid brackets.}
%  We changed the sentence to:
%  \textit{Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to various stressors
% In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism.}
%  Lines @@@

\textcolor{blue}{Line 10-11: Please rephrase to avoid brackets.}
%  We changed the sentence to:
%  \textit{We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.}
%  Lines @@@

\textcolor{blue}{Line 14-15: If space requirements permit, please include more detailed information about your results.
The current description is too vague and does not effectively convey your main findings.}
%  We've expanded a little on the results
% One point of keeping the vagueness in the abstract though was to not give the impression that we provide any prescriptive results but rather a broad understanding that the choice of metric is important.
%  \textit{We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).}
%  \textit{For example, the competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.}

\textcolor{blue}{\textbf{Introduction.}}

\textcolor{blue}{Line 25-28. It appears to me that you are using the definition of a functional trait from Violle et al.(2007) in Oikos. Please include a citation for this reference if it is the case.}
%  We've added the reference as suggested.

\textcolor{blue}{Line 37. I would avoid using the term ``new trait-focused ecology.'' While it's true that a more ``explicit'' approach to functional ecology has emerged over the past few decades, the foundation of functional analyses is rooted in the field of ecology itself.
For example, as you mentioned, Darwin's finches represent an early functional analysis centered on beak traits.
Although the specific term ``functional traits'' was less commonly used in the past, broader functional analyses were not uncommon, particularly in plant ecology (see, for example, Calow, P 1987 ``Towards a Definition of Functional Ecology.'' Functional Ecology; and Duarte et al., 1995 ``Comparative Functional Plant Ecology: Rationale and Potentials.'' Trends in Ecology \& Evolution).}
%  This is a really good point
% We've removed the mention of ``new''.

\textcolor{blue}{Line 73-74. Please clarify how this is related to the previous question ``what'' and ``how''.}
%  We've expanded the sentence to:
%  \textit{Mechanism description is then about linking the pattern and the process to some biological properties (sometimes referred to as ``rules'\,') and is at the core of answering the biological question at hand.}
%  Lines @@@

\textcolor{blue}{Line 93-96. Please explain why the chosen metric is not appropriate for the proposed example.}
%  We've expanded the sentence to:
%  \textit{In such a scenario, measuring the occupied area (i.e.~\(km^{2}\)) of each population on each island (a pattern - ``what?'\,') will not be the most appropriate way to understand the potential competition between these populations because it might not give information on whether the two populations actually interact in some way - a very likely condition for competition to take place.}
%  Lines @@@

\textcolor{blue}{Line 100. I think it would be beneficial to more clearly address your research question here. Is your goal simply to describe the patterns of functional diversity across different processes, or are you aiming to test how well different metrics capture the effects of various mechanisms? If it's the latter, how did you test this, and what are your expectations in relation to the null model?}
%  We have now clarified what our goal and expectations are in the last paragraph of the introduction:
%  \textit{Our goal is to assess the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  Lines @@@

\textcolor{blue}{Line 102. Please refer to Table 1 together with figure 1, to provide more details on the ecological mechanisms.}
%  We now refer to both the table and the figure.

\textcolor{blue}{Line 103-106.Please provide more information on why you chose to use a combination of empirical and simulated data.
From my understanding, your intention was to apply your framework within a more realistic context.
If that is the case, please clarify this point to strengthen your rationale.}
%  We changed the sentence to:
%  \textit{Beyond simulations, we also used an empirical dataset of Hawaiian bird traits and used pre-historical and historical extinction events on Hawaiian islands (process) resulting in a trait space modified by extinctions (the pattern) capturing the effects of species extinction on species trait distributions (the mechanism).}
%  Lines @@@

\textcolor{blue}{Line 106-111.
This content could be more suitable at the start of the discussion section.
Here you could maybe focus on outlining your hypotheses and predictions rather than presenting your conclusions.
By doing so, you establish a clearer context for interpreting your findings and guide the reader through the rationale behind your study.}
%  We have now moved this sentence to the end of the conclusion (Lines @@@) to end the introduction on a clear description of our goals and expectations:

% \textit{Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure \textbackslash ref\{Fig:simulations\} and \textbackslash ref\{Tab:mechanisms\}).}
%  \textit{Our goal is to test the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  \textit{To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based extinction events on Hawaiian islands before and after the year 1500 (process; how) and how different metrics (pattern; what) can lead to different interpretations of the data.}
%  \textit{We show that, that with a fixed process and mechanisms approximation, the choice of the statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data with the exact same data being sometimes interpreted in opposite ways.}

@@

\textcolor{blue}{\textbf{Methods}.}

\textcolor{blue}{Line 115. Please provide more details of the elements of your model, as not all readers may be familiar with the Brownian Motion trait model.}
%  We clarified the sentence as follows:
%  \textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e.~a pure birth speciation model) until reaching 200 elements in \textbackslash texttt\{R\} \textbackslash citep\{rcore\} using \textbackslash texttt\{treats\} \textbackslash citep\{guillerme2024treats\}.}
%  Lines @@@

\textcolor{blue}{Line 131. Please explain what the distance D is.}
%  We clarified this point as follows:
%  \textit{Increasing density: by removing the required amount of pairs of points with a variable pairwise distance of at least \(D\) (i.e.~removing the \(n\) pairs of points that are at least \(D\) distance away from each other to remove the desired number of elements).}
%  Lines @@@

\textcolor{blue}{Line 139-144. I suggest breaking this section into two sentences and improve readability.
The current structure is challenging to follow.}
%  We split this sentence in three for clarity:
%  \textit{Each bandwidth has an observed probability of sampling of \(n\) (proportional to the number of elements in that bandwidth) and is skewed to a probability of sampling of \(n 
% times m^{p}\)
% For bandwidths with a low density (low values of \(n\)) we used \(m\>0\) to increase their resampling probability, and the opposite for bandwidths with a high density (high values of \(n\); using \(m\<0\) to decrease sampling property)
% This effectively flattened the normal distribution of the data by some factor \(p\) (here we used \(p=3\)).}
%  Lines @@@

\textcolor{blue}{Table 1. Please provide references for the Equalizing and Facilitation mechanisms to offer clearer context for the focal ecological mechanisms studied.}
%  We now added references to \textbackslash citep\{chesson2000mechanisms,barot2004mechanisms\} for the equalising mechanism and to \textbackslash citep\{bruno2003inclusion, danet2024species\} for the facilitation one.
%  Lines @@@

\textcolor{blue}{Line 183. Please provide the total number of the different metrics you used and refer to the table below (table 2).}
%  We've now added a reference to table 2 and summarised the total amount of metrics.
%  Lines @@@

\textcolor{blue}{Figure 1. I believe adding a color legend to the figure would enhance its clarity.}
%  We have now reworked figure 1 to be more didactic (especially for the results interpretation)
% We have not added the color legend to the plot itself (although we have added orange arrows to make it clearer) but in the caption to not overload an already busy figure.

\textcolor{blue}{Figure 1. Please include information in the figure caption explaining what the two types of graphs represent.}
%  We added the following clarifications to the figure caption:
%  \textit{We then applied different stressors to the resulting trait space
% We represent the elements removed by the stressors in blue and the ones kept in the trait space in orange
% The first column illustrate the distribution of the elements in one dimension with the overlap between the removed (blue) and kept (orange) elements in brown (note we did not measure any metric on 1D spaces)
% The second column illustrate the same distributions in 2 dimensions
% We did not illustrate the distributions in 4 or 8 dimensions here
% We used the following stressors:}
%  Lines @@@

\textcolor{blue}{Line 194-195. Please improve clarity.}
%  @@@TODO: rewrite following other reviewers comments (stats for results)

\textcolor{blue}{Line 201-203. Please simplify and improve clarity.}
%  @@@TODO: rewrite following other reviewers comments (stats for results)

\textcolor{blue}{Line 207-211. Please provide a clearer explanation of the effect size and statistical significance of these model outcomes, similar to the explanation provided in lines 198-201.}
%  @@@TODO: rewrite following other reviewers comments (stats for results)
%  @@@TODO: calculate effect sizes
%  We now added all the statistics results in the supplementary results from table 6 to 18.

\textcolor{blue}{Line 221. Please specify the added species.}
%  @@@TODO: name of the two added marine species

\textcolor{blue}{Table 2. The aspect for the third row is missing.}
%  We are not sure about what the reviewer meant here
% The tables are all formatted in LaTeX and we prefer to leave it to the production team to set it up as they like should the manuscript be accepted.

\textcolor{blue}{\textbf{Results}.}

\textcolor{blue}{Line 248-251. Please provide more detail regarding these results (e.g., what the scaled metric scores are, the effect size, and the significance of data reduction caused by the stressor). Additionally, refer to the relevant section of Figure 2 (to explain how the observed trends in the figure support this assertion), and to the tables in supplementary material.}
%  @@@TODO: add stats for results (following other reviewers comments)

\textcolor{blue}{Line 252. I would avoid the term ``naively''.}
%  We removed this term.

\textcolor{blue}{Line 254. Please ensure that the use of italics follows the guidelines from the journal.}
%  We did not find any specific guideline regarding emphasising words with italics but removed it nonetheless.

\textcolor{blue}{Line 254-256. Apply the same changes mentioned in the comment below to lines 248-251.}
%  @@@TODO: add stats for results (following other reviewers comments)

\textcolor{blue}{Line 260. Please avoid using the exclamation mark.}
%  We changed the punctuation to a dot.

\textcolor{blue}{Line 262-269. Please provide more detail regarding these results. Additionally, refer to the table in supplementary material.}
%  @@@TODO: add stats for results (following other reviewers comments)
%  We also now refer to the supplementary results:
%  \textit{For the empirical data, we see again a range of different possible interpretations depending on the metric describing the pattern (Figure \textbackslash ref\{Fig:empirical\_results\}, Supplementary table 5).}
%  Lines @@@

\textcolor{blue}{Line 275-283. This should be in the discussion section.}
%  We moved this paragraph to the discussion on lines @@@.

\textcolor{blue}{Figure 2. Maybe I misunderstood, but I think that you are missing the adjusted R square from the figure.}

%  We removed mention of the R2 (from a previous cluttered version of the figure) and now refer to it in the supplementaries in the figure caption.

\textcolor{blue}{Figure 3. Please provide more details in the figure caption. Please include the results from the t-test in the supplementary material.
While the significance of the test is demonstrated through the gray plots in Figures 2 and 3, the relative values are also necessary.}
%  The tables in the supplementary materials now contain the R2 values for both the Figures 2 and 3
% We have now clarified that in the figure captions:
%  \textit{The model results (including \(R^{2}\) and t-statistics) are available in supplementary table 4.}
%  Lines @@@ and Lines @@@

\textcolor{blue}{\textbf{Discussion}.}
\textcolor{blue}{Line 285-291. This section effectively summarizes your main goal and key message.
However, a more detailed interpretation of your main findings is needed.
I believe that simply stating that ``different metrics capture different patterns (what), leading to inferences about different processes (how) and mechanisms (why)'' is insufficient.
I recommend focusing more on the performance of the various metrics to offer readers a clearer framework for selecting the appropriate metric for each mechanism.
Please refer to items 3 and 4 for further details.}

\textcolor{blue}{Line 291-302.
I appreciate the detailed discussion of the potential limitations of your methodology.
However, in my opinion, this should follow a more in-depth interpretation of your results and their relation to the ``what'', ``how'', and ``why'' questions.
Consider placing the discussion of caveats and limitations in a separate paragraph, where you address the constraints of your framework.}

\textcolor{blue}{Line 303-315.
Based on your preference, I recommend moving this section further into the discussion.
This will allow you to address the importance of selecting the appropriate metric in relation to the number of dimensions used for the trait space.}
% We have now re-orginased the discussion and clearly separated each subsection
% The first one now discusses the results (simulated and empirical), the second one focuses on the caveats of our approach.
% And the third section is now a standalone box (box 2) discussing how we believe the what, how and why questions can help workers choose their functional diversity metric of interest based on their own data and question
% Hopefully these clear distinctions will make it easier to follow (or skip) for the reader.

\textcolor{blue}{Line 313-316. Please rephrase to improve readability.}
%  We have now removed this sentence.

\textcolor{blue}{Line 316-337. I think this section should be simplified.
The discussion focuses too heavily on the limitations of the approach, making it overly methodological.
A greater emphasis on the ecological implications of the results would better convey your main message.
I recommend streamlining this section and, if preferred, moving part of it to the supplementary material, where you can address the implications of scaling your results with the null model.}
%  We have now moved this section into the named ``caveats'' section
% Hopefully this will make it easier to skip for readers no interested in the methodological aspects of our work
% However, if the reviewer still thinks this section is too focused, we're happy to reduce it in a future version.

\textcolor{blue}{Line 321-324. When you refer to your findings, please provide a more explicit reference to the values you found and the relative graph to better guide the reader through your results and your interpretation.}
%  We have now specified which part of figure 2 we were using to illustrate this point:
%  \textit{(e.g.~facilitation mechanism measured as divergence using the probability density method - not distinguishable from null when removing 80\textbackslash\% or 60\textbackslash\% of the data; Figure \textbackslash ref\{Fig:simulations\})}
%  Lines @@@

\textcolor{blue}{Line 327. Please replace `workers' with `ecologists,' as you are referring to an example from an ecological study.}
%  We changed ``workers'' to ``researchers'' because not only ecologists are likely to work on such data/questions (e.g.~evolutionary biologists, palaeobiologists, etc.)

\textcolor{blue}{Line 339-340. Please simplify and increase readability.}
%  We reworded this sentence as follows:
%  \textit{The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity since 1500 is not greater than expected by chance (at least for certain metrics; Figure \textbackslash ref\{Fig:empirical\_results\}) because one would expect that extinctions since 1500 target specific areas of functional diversity space (i.e.~birds with specific trait combinations).}
%  Lines @@@
